### 2020-07-15 code/figure1.Rmd

```{r fig-1a-1l, warning=FALSE, fig.height=11, fig.width=8.5}
load("../data/cancer_disrupt_heatmap.Rda")
umap_cancertype <- fread("../data/umap_2d_3d_coors.tsv")

@@

## violin mutation count
fig1G <- ggplot(umap_cancertype, aes(x = knn_clust, y = mut_count)) +
    geom_violin(aes(fill = knn_clust)) + #draw_quantiles = c(0.25, 0.5, 0.75)) 
    geom_boxplot(aes(fill = knn_clust), width=0.2) +
    theme_classic() +
    scale_y_log10() +
    xlab("") +
    ylab("Log(Mutation Count)") +
    labs(color = "", tag = "G") +
    theme(axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 12),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_text(size = 12),
          legend.position = "none") +
  scale_fill_manual(values = kelly.colours[-c(1:2)])

@@

## umap brca somatic mutations
load("../data/gdc_gene_exp_adj.Rda")
load("../data/gene_data.Rda")
brca <- gene_data %>% .[.$external_gene_name %in% c("BRCA1", "BRCA2"),] %>% dplyr::select(ensembl_gene_id, external_gene_name) ## Getting BRCA gene information
brca <- gdc_gene_exp_adj %>% .[rownames(.) %in% brca$ensembl_gene_id, ] %>% ## we only want brca gene
  rownames_to_column(var = "ensembl_gene_id") %>%
  left_join(., brca, by = "ensembl_gene_id") %>% 
  remove_rownames()  %>% column_to_rownames(var = "external_gene_name") %>% ## replace ensembl_id with gene_name
  dplyr::select(-ensembl_gene_id) %>% ## get ride of ensembl_id  col
  t() %>% as.data.frame() %>% ## correct orientation for umap_cancertype df
  mutate(BRCA1 = case_when(BRCA1 == "1" ~ "BRCA1",
                           BRCA1 == "0" ~ "NA"),
         BRCA2 = case_when(BRCA2 == "1" ~ "BRCA2",
                           BRCA2 == "0" ~ "NA")) %>%
  mutate(mut_type = gsub('NA,|,NA', '', paste(.$BRCA1, .$BRCA2, sep = ","))) %>% ## combine mutation cause some patients can have both
  dplyr::select(-BRCA1, -BRCA2) %>% ## remove col we don't need
  rownames_to_column(var = "V1") %>% ## getting data ready for leftjoin 
  left_join(., umap_cancertype[,c("V1", "x_2d", "y_2d", "project_code")], by = "V1") ## leftjoin by sample id

@@

layout <- '
AAAABBCCDD
AAAAEEEFFF
AAAAGGGG##
AAAAGGGG##
AAAAHHIIJJ
AAAAKKLL##
'

@@

```{r mutation-burden-plot, message = FALSE}
umap_cancertype <- fread("../data/umap_2d_3d_coors.tsv")

mut_order <- list() ## store data in list 
for(n in 1:max(umap_cancertype$knn_clust)) { ## loop through all the clusters
  message(n)
  mut_order[[n]] <- umap_cancertype %>% filter(knn_clust == n) %>%  ## filter for specific class
    dplyr::select(V1, x_2d, y_2d, knn_clust, mut_count) %>% ## select relevant columns
    arrange(mut_count) ## arrange mutation in increasing order
}
mut_order <- mut_order %>% do.call(rbind, .) %>% as.data.frame() ## rbind list

total_n <- as.data.frame(table(mut_order$knn_clust))

## for making dup discrete x-axis
guide_axis_label_trans <- function(label_trans = identity, ...) {
  axis_guide <- guide_axis(...)
  axis_guide$label_trans <- rlang::as_function(label_trans)
  class(axis_guide) <- c("guide_axis_trans", class(axis_guide))
  axis_guide
}

guide_train.guide_axis_trans <- function(x, ...) {
  trained <- NextMethod()
  trained$key$.label <- x$label_trans(trained$key$.label)
  trained
}

ggplot(mut_order, aes(x = V1, y = log10(mut_count + 1), color = knn_clust)) + 
  geom_point() + 
  ylab("Log10(Mutation Count)") + 
  xlab("") +
  scale_x_discrete(breaks = c(as.character(mut_order$V1[300]), as.character(xaxis_breaks$V1)), labels = c("Class =" ,1:12)) +
  guides(x.sec = guide_axis_label_trans(~paste(c("n = ", total_n$Freq)))) + ## create second x-axis
  scale_y_continuous(sec.axis = sec_axis(~.)) + ## dup y-axis
  #scale_y_log10() +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        legend.position = "none") +
  scale_color_manual(values = kelly.colours[-c(1:2)])
```

>>>>

```{r fig-1a-1l, warning=FALSE, fig.height=11, fig.width=8.5, message = FALSE}
load("../data/cancer_disrupt_heatmap.Rda")
umap_cancertype <- fread("../data/umap_2d_3d_coors.tsv")

@@

## mutation burden plot
mut_order <- list() ## store data in list 
for(n in 1:max(as.numeric(as.character(umap_cancertype$knn_clust)))) { ## loop through all the clusters
  message(n)
  mut_order[[n]] <- umap_cancertype %>% filter(knn_clust == n) %>%  ## filter for specific class
    dplyr::select(V1, x_2d, y_2d, knn_clust, mut_count) %>% ## select relevant columns
    arrange(mut_count) ## arrange mutation in increasing order
}
mut_order <- mut_order %>% do.call(rbind, .) %>% as.data.frame() ## rbind list

total_n <- as.data.frame(table(mut_order$knn_clust)) ## getting total n in cluster

mut_order$log10mut <- log10(mut_order$mut_count + 1) ## mutation count log transformation + pseudo count

mut_median <- mut_order %>% 
  group_by(knn_clust) %>% 
  dplyr::summarize(mut_med = median(log10mut)) ## calculate median for each class

fig1G <- ggplot(mut_order, aes(x = V1, y = log10mut, color = knn_clust)) + 
  geom_point() + 
  ylab("Log10(Mutation Count)") + 
  xlab("") +
  theme_classic() +
  facet_wrap(~ factor(knn_clust, levels = c(1:12)), ncol = 12, scales = "free_x") +
  geom_hline(data= mut_median, aes(yintercept=mut_med)) +
  scale_y_continuous(sec.axis = sec_axis(~.)) + ## dup y-axis
  scale_x_discrete(breaks = c(as.character(mut_order$V1[300]), as.character(xaxis_breaks$V1)), labels = c("n =" ,total_n$Freq)) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.ticks.x = element_blank(),
        axis.line.x = element_line(colour = "black", size = 0.5),
        axis.line.x.top = element_line(colour = "black", size = 0.5),
        #strip.background.x = element_blank(),
        strip.placement = "outside",
        strip.text.x = element_text(size = 12),
        legend.position = "none") +
  scale_color_manual(values = kelly.colours[-c(1:2)])

## code to keep only the bottom border line of the facet
fig1G <- ggplotGrob(fig1G)
lg <- linesGrob(x=unit(c(0,1),"npc"), y=unit(c(0,0),"npc"), 
                   gp=gpar(col="black", lwd=4))

for (k in grep("strip-t",fig1G$layout$name)) {
  fig1G$grobs[[k]]$grobs[[1]]$children[[1]] <- lg
}

fig1G <- as.ggplot(fig1G)

@@

## umap brca somatic mutations
load("../data/gdc_gene_exp_adj.Rda")
load("../data/gene_data.Rda")
umap_cancertype <- fread("../data/umap_2d_3d_coors.tsv")
brca <- gene_data %>% .[.$external_gene_name %in% c("BRCA1", "BRCA2"),] %>% dplyr::select(ensembl_gene_id, external_gene_name) ## Getting BRCA gene information
brca <- gdc_gene_exp_adj %>% .[rownames(.) %in% brca$ensembl_gene_id, ] %>% ## we only want brca gene
  rownames_to_column(var = "ensembl_gene_id") %>%
  left_join(., brca, by = "ensembl_gene_id") %>% 
  remove_rownames()  %>% column_to_rownames(var = "external_gene_name") %>% ## replace ensembl_id with gene_name
  dplyr::select(-ensembl_gene_id) %>% ## get ride of ensembl_id  col
  t() %>% as.data.frame() %>% ## correct orientation for umap_cancertype df
  rownames_to_column(var = "V1") %>% ## getting data ready for leftjoin 
  mutate(BRCA1 = case_when(BRCA1 == "1" ~ "BRCA1",
                           BRCA1 == "0" ~ "NA"),
         BRCA2 = case_when(BRCA2 == "1" ~ "BRCA2",
                           BRCA2 == "0" ~ "NA")) %>%
  mutate(mut_type = gsub('NA,|,NA', '', paste(.$BRCA1, .$BRCA2, sep = ","))) %>% ## combine mutation cause some patients can have both
  dplyr::select(-BRCA1, -BRCA2) %>% ## remove col we don't need
  left_join(., umap_cancertype[,c("V1", "x_2d", "y_2d", "project_code")], by = "V1") ## leftjoin by sample id

@@

layout <- '
AAAABBCCDD
AAAAEEEFFF
AAAAGGGGGG
AAAAGGGGGG
AAAAHHIIJJ
AAAAKKLL##
'

@@

### 2020-07-15 code/figure4.Rmd

```{r figure-4, message = FALSE, fig.cap="**Figure 4: Some genes with multiple pathway memberships drive clustering.** Each column is a tumor in class 6 or 7, each row represents a gene with mutation (red) or not mutated (grey). The genes are organized according to which of the top distinguishing pathways for this class that they effect. Some genes affect multiple pathways, as indicated by the bars at left. The highlighted genes at right are all proteasomal subunits.", eval=FALSE}

load("../data/class67_gene_2_path_heatmap.Rda") ## load data
load("../data/path_match_in_set.Rda")

>>>>

```{r figure-4, message = FALSE, fig.cap="**Figure 4: Some genes with multiple pathway memberships drive clustering.** Each column is a tumor in class 6 or 7, each row represents a gene with mutation (red) or not mutated (grey). The genes are organized according to which of the top distinguishing pathways for this class that they effect. Some genes affect multiple pathways, as indicated by the bars at left. The highlighted genes at right are all proteasomal subunits.", eval=FALSE}
load("../data/class67_gene_2_path_heatmap.Rda") ## load data
load("../data/path_match_in_set.Rda")
